Lesson 4
========================================================
```{r}
getwd()
list.files()

pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
```


***

### Scatterplots and Perceived Audience Size
Notes: best for 2 continuous variables

***

### Scatterplots
Notes: lie about their age

```{r Scatterplots}
qplot(x = age, y = friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) + geom_point() + xlim(13, 90)

summary(pf$age)
```

***

### Overplotting
Notes: too many points, jitter makes the point more blurred so the bar-like effects are 
gone

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_jitter(alpha = 1/20) + xlim(13, 90)

# alpha = 1/20 means take 20 points as 1, i.e. takes 20 points for a circle to appear dark

```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20) + 
  xlim(13, 90) + coord_trans(y = 'sqrt')

# more complex syntax because don't want to get negative friend count with the jitter
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) + 
  xlim(13, 90) + coord_trans(y = 'sqrt')

# more complex syntax because don't want to get negative friend count with the jitter
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) + 
  coord_trans(y = 'sqrt')

# really large values, so use sqrt transformation again
```

***

### Overplotting and Domain Knowledge
Notes: % of audience as a transformation, a specific post (one post) vs. general

***

### Conditional Means
Notes: Common functions: filter(), group_by(), mutate(), arrange()

```{r Conditional Means}
install.packages('dplyr') # NEED TO INSTALL THIS
library(dplyr)

age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups,
                friend_count_mean = mean(friend_count),
                friend_count_median = median(friend_count),
                n = n())

pf.fc_by_age <- arrange(pf.fc_by_age, age)

head(pf.fc_by_age)

pf.fc_by_age_2 <- pf %.%
  group_by(age) %.%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %.%
  arrange(age)

head(pf.fc_by_age_2)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(age, friend_count_mean), data = pf.fc_by_age) +
  geom_line()

# can immediate see the oddness of age 69
# highs and lows at higher age
```

***

### Overlaying Summaries with Raw Data
Notes: good stuff overlaying summaries

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0), color = 'orange') + 
  coord_cartesian(xlim = c(13, 70), ylim = c(0,1000)) + xlim(13, 90) +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, probs = 0.1, linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, probs = 0.9, linetype = 2, color = 'blue') + 
  geom_line(stat = 'summary', fun.y = quantile, probs = 0.5, color = 'blue')

```

#### What are some of your observations of the plot?
Response: More than 1000 friends are rare, people between 30 and above have less than 250 friends

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count, method = 'pearson')

with(pf, cor.test(age, friend_count, method = 'pearson')) 
# create an environment with the data with the 'with' function
# wouldn't want to include older ages in the number because they're likely to be 
# incorrect

# no meaningful correlation
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes: single number coefficients are useful but scatterplots and visualizations are 
important

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(y = likes_received, x = www_likes_received), data = pf) +
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$www_likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'red')

# lm = linear model, smoother
```

***

### Strong Correlations
Notes: strong correlation because 1 is a superset of the other

```{r Strong Correlations}
with(pf, cor.test(www_likes_received, likes_received))

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes: number of status updates per month are highly correlated with how many days in the last month the logged in, how many friends they have, how many photos they've uploaded. because they all measure the same thing --> engagement. correlation helpful in linear model determination

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)
?Mitchell
```

Create your plot!

```{r Temp vs Month}
qplot(data = Mitchell, x = Month, y = Temp)

ggplot(aes(x = Month, y = Temp), data = Mitchell) + geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x = Month, y = Temp), data = Mitchell) + geom_point() + geom_line() +
  scale_x_discrete(breaks=seq(0,203,12))

ggplot(aes(x=(Month%%12),y=Temp),data=Mitchell)+ 
  geom_point()

# example dataset

install.packages("energy")
library(energy)

x <- seq(0, 4*pi, pi/20)
y <- cos(x)
qplot(x = x, y = y)
dcor.ttest(x, y)
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + geom_line()
```

***

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- (12 - pf$dob_month)/12 + pf$age

```

Programming Assignment
```{r Programming Assignment}
age_groups_month <- group_by(pf, age_with_months)
pf.fc_by_age_months <- summarise(age_groups_month,
                friend_count_mean = mean(friend_count),
                friend_count_median = median(friend_count),
                n = n())

pf.fc_by_age_months <- arrange(pf.fc_by_age_months, age_with_months)

head(pf.fc_by_age_months)
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(age_with_months, friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months <= 71)) +
  geom_line()
```

***

### Smoothing Conditional Means
Notes: systematic errors in model generating

```{r Smoothing Conditional Means}
ggplot(aes(age_with_months, friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months <= 71)) +
  geom_line() + geom_smooth()
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection: create multiple visualizations of the same data, reveal different things

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

