Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
getwd()
setwd('/Users/Warren/Desktop/Udacity/EDA_Course_Materials/lesson3')
list.files()

pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)

names(facebook)
qplot(x = dob_day, data = facebook) + scale_x_discrete(breaks=1:31)
pf <- facebook
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes: 
People's perception of audience on facebook match up to reality (business: because 
it would effect the way you present yourself)
Survey people how many people do you think saw this post (big mismatch of people's 
perceived and actual audience size)

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes: people underestimate their audience

***
### Faceting
Notes:

```{r Faceting}
qplot(x = dob_day, data = pf) + scale_x_discrete(breaks=1:31) +
facet_wrap(~dob_month, ncol = 3)
```

#### Let’s take another look at our plot. What stands out to you here?
Response: default setting jan 1st

***

### Be Skeptical - Outliers and Anomalies
Notes: maybe true or false or limitation of data

***

### Moira's Outlier
Notes: see the bulk of data
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x = friend_count, data = pf)
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(x = friend_count, data = pf, xlim = c(0,1000))

qplot(x = friend_count, data = pf) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))


```

### Exploring with Bin Width
Notes: people respond in whole numbers e.g. 50, 100

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) + facet_wrap(~gender)
```

***

### Omitting NA Values
Notes: na.omit omits everything that has null values

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) + facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}
qplot(x = tenure, data = pf, binwidth = 30, color = I('black'), 
      fill = I('#099009'))
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(x = tenure/360, data = pf, binwidth = .25, color = I('black'),
      xlab = 'Number of years using Facebook', 
      ylab = 'Number of users in sample',
      fill = I('#F79420')) + scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0,7))
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}

```

***

### User Ages
Notes:

```{r User Ages}
qplot(x = age, data = subset(pf, !is.na(pf$gender)), color = I('black'), binwidth = 1,
      xlab = 'Age of users in Facebook', 
      ylab = 'Number of users in sample',
      fill = I('#F79420')) + scale_x_discrete(breaks = seq(1, 120, 30)) +
  facet_wrap(~gender)

by(pf$age, pf$gender, summary)

# female users are in general older

```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes: regression assumes normality of variable

```{r}

install.packages('grid.Extra')
library(gridExtra)

p1 <- qplot(x = friend_count, data = pf)
p2 <- qplot(x = log10(friend_count + 1), data = pf)
p3 <- qplot(x = sqrt(friend_count), data = pf)

grid.arrange(p1, p2, p3, ncol = 1)



```

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
p1 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p2 <- p1 + scale_x_log10()
p3 <- p2 + scale_x_sqrt()

```

***


### Frequency Polygons

```{r Frequency Polygons}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10, geom = 'freqpoly',
      color = gender) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))


qplot(x = friend_count, y = ..count../sum(..count..),
      xlab = 'Friend Count', ylab = 'Proportion of Users with that friend count',
        data = subset(pf, !is.na(gender)), binwidth = 10, geom = 'freqpoly',
      color = gender) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))


```

***

### Likes on the Web
Notes: female likes 2 times more than male. can inform a website which feature to keep
and which to leave behind

```{r Likes on the Web}
qplot(x = www_likes, data = subset(pf, !is.na(gender)), geom = 'freqpoly',
      color = gender) + scale_x_continuous() + scale_x_log10()
```


***

### Box Plots
Notes:

```{r Box Plots}
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot',
      color = gender)
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot',
      color = gender, ylim = c(0,1000))

qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot',
      color = gender) + scale_y_continuous(limits=c(0,1000))
```

***

### Box Plots, Quartiles, and Friendships
Notes: female more friends than male (median)

```{r Box Plots, Quartiles, and Friendships}
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot',
      color = gender) + coord_cartesian(ylim=c(0,250))

by(pf$friend_count, pf$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response: Median of female is higher than male
#### Write about some ways that you can verify your answer.
Response: Boxplots are much richer than numeric summaries with information
```{r Friend Requests by Gender}
qplot(x = gender, y = friendships_initiated, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot',
      color = gender) + coord_cartesian(ylim = c(0,150))

by(pf$friendships_initiated, pf$gender, summary)
```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
summary(pf$mobile_likes)

summary(pf$mobile_likes > 0)

mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

sum(pf$mobile_check_in == 1)/ length(pf$mobile_check_in)
```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!