Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
getwd()
setwd('/Users/Warren/Desktop/Udacity/EDA_Course_Materials/lesson3')
list.files()

pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape =  4)

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color = gender), stat = 'summary', fun.y = median) 
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  ungroup() %>%
  arrange(age)

ggplot(aes(x = age, y = friend_count_mean, color = gender), data = pf.fc_by_age_gender) +
  geom_line()

```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'friend_count_median')

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
pf.fc_by_age_gender.wide$ratio <- 
  pf.fc_by_age_gender.wide$female / pf.fc_by_age_gender.wide$male

ggplot(aes(x = age, y = ratio), data = pf.fc_by_age_gender.wide) + geom_line() + 
  geom_hline(aes(yintercept = 1), linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- 2014 - ceiling(pf$tenure/365)
head(floor(2014 - pf$tenure/365))
head(pf$year_joined)

table(pf$year_joined)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, 
                             breaks = c(2004, 2009, 2011, 2012, 2014))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count),
       data = pf) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median) 
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = pf) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)

```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure >= 1), summary(friend_count / tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) + 
  geom_line(aes(color = year_joined.bucket)) 

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
getwd()
setwd('/Users/Warren/Desktop/Udacity/EDA_Course_Materials/lesson5')
list.files()

yo <- read.csv('yogurt.csv')
str(yo)

# change id to factor
yo$id <- factor(yo$id)
str(yo)

# histogram of prices
qplot(price, data = yo, binwidth = 10)
table(yo$price)

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
length(unique(yo$price))

yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain +
            mixed.berry)

qplot(data = yo, x = all.purchases)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x = time, y = price), data = yo) +
  geom_jitter(alpha = 1/4, shape = 21)
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)

sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price),
       data = subset(yo, id %in% sample.ids)) +
  facet_wrap(~id) +
  geom_line() +
  geom_point(aes(size = all.purchases), pch = 1)


set.seed(4220)

sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price),
       data = subset(yo, id %in% sample.ids)) +
  facet_wrap(~id) +
  geom_line() +
  geom_point(aes(size = all.purchases), pch = 1)


```

***

### The Limits of Cross Sectional Data
Notes: cross-section vs. time-series

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

```{r}
install.packages("GGally")
library(GGally)

theme_set(theme_minimal(20))

set.seed(1836)
pf_subset <- pf[, c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])



```
***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
list.files()
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
library(reshape2)
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

